---
title: "John's Titanic @ Kaggle Notebook"
output: html_notebook
---

Let's read in the training data and take quick look

```{r}
train_data <- read.table("train.csv",header=TRUE,sep=",",na.strings="")
str(train_data)
```

Here is a variable glossary taken from elsewhere

Variable Name | Description
--------------|-------------
Survived      | Survived (1) or died (0)
Pclass        | Passenger's class
Name          | Passenger's name
Sex           | Passenger's sex
Age           | Passenger's age
SibSp         | Number of siblings/spouses aboard
Parch         | Number of parents/children aboard
Ticket        | Ticket number
Fare          | Fare
Cabin         | Cabin
Embarked      | Port of embarkation

## Feature Engineering

I'm following an example. In that example they grab more features from the full name.

### Title 

```{r}
# Grab title from passenger names
train_data$Title <- gsub('(.*, )|(\\..*)', '', train_data$Name)

# Show title counts by sex (before)
table(train_data$Sex, train_data$Title)

# Clean up titles
train_data$Title[train_data$Title == 'Mlle'] <- 'Miss' 
train_data$Title[train_data$Title == 'Mme'] <- 'Miss' 
train_data$Title[train_data$Title == 'Ms'] <- 'Miss' 
#
train_data$Title[train_data$Title == 'Capt'] <- 'Officer' 
train_data$Title[train_data$Title == 'Col'] <- 'Officer' 
train_data$Title[train_data$Title == 'Major'] <- 'Officer' 
#
train_data$Title[train_data$Title == 'Don'] <- 'Royal' 
train_data$Title[train_data$Title == 'Lady'] <- 'Royal' 
train_data$Title[train_data$Title == 'Sir'] <- 'Royal' 
train_data$Title[train_data$Title == 'the Countess'] <- 'Royal' 
#
train_data$Title[train_data$Title == 'Dr'] <- 'Professional' 
train_data$Title[train_data$Title == 'Jonkheer'] <- 'Professional' 
train_data$Title[train_data$Title == 'Rev'] <- 'Professional' 

# Show title counts by sex (after)
table(train_data$Sex, train_data$Title)

```

### Surname

```{r}
# function to get surname
get_surname <- function(x) {
  strsplit(as.character(x),"[,.]")[[1]][1]
}

# populate new variable
train_data$Surname <- sapply(train_data$Name,get_surname)

# sample results
table(train_data[seq(1,25),c("Surname")],useNA="always")

# make sure no null's
length(train_data[is.na(train_data$Surname)])
```

### Family

Again following example we calculate a family size feature and another variable which is surname with family size

```{r}
#calculate family size
train_data$FamilySize <- train_data$SibSp + train_data$Parch + 1
table(train_data$FamilySize)
#build family name + size variable
train_data$FamilyInfo <- paste(train_data$Surname,train_data$FamilySize,sep="_")
#sample results
train_data[seq(1,10),c("Surname","FamilySize","FamilyInfo")]
```

Let's plot this data

```{r, message=FALSE, warning=FALSE}
#plot familly size and surivale details
library(ggplot2)
ggplot(train_data, aes(x = FamilySize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') 
```

Still following the example we leverage chart above to see some categories or discrete levels which may be helpful to instantiate

```{r}
#discretize family size
train_data$FamilySizeClass[train_data$FamilySize == 1] <- 'singleton'
train_data$FamilySizeClass[
  train_data$FamilySize < 5 & train_data$FamilySize> 1
] <- 'small'
train_data$FamilySizeClass[train_data$FamilySize > 4] <- 'large'

#show family size by survival using a mosaic plot
mosaicplot(
  table(train_data$FamilySizeClass, train_data$Survived), 
  main='Family Size by Survival',
  shade=TRUE
)

#similar plot with ggplot
ggplot(data=train_data) + stat_count(aes(FamilySizeClass,fill=factor(Survived)),position="dodge")
```

### Cabin Class

The example talks about cabin class, I had a prior implementation will use

```{r}
train_data$CabinClass <- as.factor(
  ifelse(
    grepl(" ",train_data$Cabin,fixed=TRUE),
    paste(substr(train_data$Cabin,1,1),substr(train_data$Cabin,1,1),sep=""),
    substr(train_data$Cabin,1,1)
  )
)
table(train_data$CabinClass)
```

## Missing Data

Following the example here. This is interesting to me as I haven't seen this technique before.

```{r}
asdf
```